Stochastic Variance Reduced Policy Gradient

### 2018-01-23 svrpg.tex
finally done writing the lemmas

%%%USEFUL PACKAGES%%%
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{amsmath,amssymb}
\usepackage{mathtools}
\usepackage{pifont}
%%Theorems
\usepackage{amsthm}
\theoremstyle{remark}
\newtheorem*{rem}{Remark}
\newtheorem*{note}{Note}
\newtheorem{case}{Case}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{conj}{Conjecture}
\newtheorem{example}{Example}
%Restatable
\usepackage{thmtools, thm-restate}
\declaretheorem[numberwithin=section]{thm}
\declaretheorem[sibling=thm]{lemma}
\declaretheorem[sibling=thm]{corollary}
\declaretheorem[sibling=thm]{assumption}
\declaretheorem[sibling=thm]{theorem}
%e.g. ...
\usepackage{xspace}
\DeclareRobustCommand{\eg}{e.g.,\@\xspace}
\DeclareRobustCommand{\ie}{i.e.,\@\xspace}
\DeclareRobustCommand{\wrt}{w.r.t.\@\xspace}
% graphics
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{tikz,pgfplots}
\pgfplotsset{compat=newest}
%% TODOs
% \setlength{\marginparwidth}{2cm}
\usepackage[colorinlistoftodos, textwidth=20mm]{todonotes}
\definecolor{citrine}{rgb}{0.89, 0.82, 0.04}
\definecolor{blued}{RGB}{70,197,221}
%%todo by Marcello
\newcommand{\todomarc}[1]{\todo[color=green, inline]{\small #1}}
\newcommand{\todomarcout}[1]{\todo[color=green]{\scriptsize #1}}
%%todo by Matteo Papini
\newcommand{\todomat}[1]{\todo[color=citrine, inline]{\small #1}}
\newcommand{\todomatout}[1]{\todo[color=citrine]{\scriptsize #1}}
%%todo by Matteo Pirotta
\newcommand{\todopir}[1]{\todo[color=blued, inline]{\small #1}}
\newcommand{\todopirout}[1]{\todo[color=blued]{\scriptsize #1}}
%%%%%%

@@

\begin{abstract}
Why we need variance-reduced gradient in RL?
\end{abstract}

\section{Introduction}
\begin{itemize}
        \item Similarities between RL and SL
        \item SL needs SVRG so RL
        \item Potential advantages of SVRG
        \item Difficulties in RL \wrt SL
\end{itemize}

\section{Preliminaries}
In this section we provide essential background on policy gradient methods for Reinforcement Learning and stochastic variance-reduced gradient methods for finite-sum optimization. In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.

\subsection{Policy Gradient}
A Reinforcement Learning task \cite{sutton2011reinforcement} can be modeled with a discrete-time continuous Markov Decision Process (MDP) $\langle\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\rangle$, where $\Sspace$ is a continuous state space; $\Aspace$ is a continuous action space; $\Transition$ is a Markovian transition model, where $\Transition(s'|s,a)$ defines the transition density from state $s$ to $s'$ under action $a$; $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$ and $R$ is the maximum absolute-value reward; $\gamma\in[0,1)$ is the discount factor; and $\rho$ is the initial state distribution.
The agent's behavior is modeled as a policy $\pi$, where $\pi(\cdot|s)$ is the density distribution over $\Aspace$ in state $s$.
% We consider episodic tasks, \ie tasks composed of episodes of length $H$, also called time horizon.
% In this setting, a trajectory $\tau$ is the sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following some stationary policy in a given episode, where $s_0$ is always sampled from the initial state distribution $\rho$.
We consider episodic MDPs with effective horizon $H$, \ie optimal policy is able to reach the target state (\ie absorbing state) in at most $H$ steps (the true length depend on the initial state). This has not to be confused with a finite horizon problem where the optimal policy is non-stationary. In this setting, we can limit our attention to trajectories of length $H$. A trajectory $\tau$ is a sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following a stationary policy, where $s_0 \sim \rho$.
With abuse of notation, we denote with $\pi(\tau)$ the density distribution induced by policy $\pi$ on the set $\Tspace$ of all possible trajectories, and with $\Reward(\tau)$ the total discounted reward provided by trajectory $\tau$:
\begin{align*}
\Reward(\tau) = \sum_{t=0}^{H-1}\gamma^t\Reward(s_t,a_t).
\end{align*}
Policies can be ranked based on their expected total reward:
\begin{align*}
	J(\pi) = \int_{\Tspace}\pi(\tau)\Reward(\tau)\de \tau.
\end{align*}
Solving an MDP means finding $\pi^* \in \argmax_{\pi} \{J(\pi)\}$.

Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable \wrt the $m$-dimensional parameter vector $\vtheta$. We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $\pol(\tau)$ with $\pi(\tau|\vtheta)$ for the sake of readability. The search for a locally optimal policy is performed by means of gradient ascent, where the gradient of $J(\vtheta)$ \wrt the policy parameters is \cite{sutton2000policy}:
\todopir{stochastic or mini-batch gradient}
\begin{align*}
\gradJ{\vtheta} = \int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
\end{align*}
We consider the on-line learning problem, where trajectories must be sampled directly by interacting with the environment. In this setting, stochastic gradient ascent is typically employed: after sampling a trajectory $\tau$ or a batch of $N$ trajectories $\Dataset_{N}$,
policy parameters are updated as $\vtheta\gets\vtheta + \alpha\gradApp{\vtheta}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of the true gradient. The REINFORCE gradient estimator \cite{williams1992simple} provides a simple, unbiased way of estimating the gradient:
\begin{align*}
\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\left(\sum_{h=0}^{H-1}\score{}{a_h^n\mid s_h^n}\right)\left(\sum_{h=0}^{H-1}\gamma^h r_h^n - b\right),
\end{align*}
where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and b can be any baseline, provided it is constant \wrt to actions.
The G(PO)MDP gradient estimator \cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance while preserving the unbiasedness:
\begin{align*}
\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\gamma^h r_h^n - b\right).
\end{align*}

@@

\section{Convergence Guarantees}

\begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
Provided $J(\vtheta)$ is L-smooth with Lipschitz constant $L$ and step sizes are sufficiently small, the parameter vector $\vtheta_A$ returned by Algorithm \ref{alg:svrpg} after $T=m\times S$ iterations has, for some constants $\psi>0, D>0$ and $c_{t}>0$ for $t=1,\dots,m$, the following property:
\[
	\EVV[]
	{\norm[]{\nabla J(\vtheta_A)}^2} \leq
		\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T} +
		\frac{Z}{N},
\]
where
\begin{align*}
	Z &= D\max_{s}\EVV[]{\norm[]{\gradApp{\tilde{\vtheta^s}}{1}}^2} \\
	&+ \max_{s,t}\left|\Tr\Cov\left[
		\gradApp{\tilde{\vtheta^s}}{1},
		c_{t+1}\vtheta_t^{s+1} + \alpha_t\nabla J(\vtheta_t^{s+1})\right]\right|
\end{align*}
\end{restatable}

@@

%Lemma 2
\begin{restatable}[]{lemma}{auxtwo}\label{lemma:aux2}
The expected squared norm of the SVRPG gradient can be bounded as follows:
\[
\Ets{\gradBlack{\vtheta_t^{s+1}}} \leq
\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
\nonumber 
+dG^2R^2\left(M_2-1\right)
\]
\end{restatable}
\begin{proof}

	\begin{align}
	\Ets{\gradBlack{\vtheta_t^{s+1}}} 
	&= \Ets{\norm[]{\gradApp{\tilde{\vtheta}^s}{N}
			+ \score{\vtheta_t^{s+1}}{\tau_i} -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}^2} \nonumber\\
	&= \Ets{\norm[]{\gradApp{\tilde{\vtheta}^s}{N}
			+ \score{\vtheta_t^{s+1}}{\tau_i} -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}-\gradJ{\vtheta_t^{s+1}} + \gradJ{\tilde{\vtheta}^s}
			+\gradJ{\vtheta_t^{s+1}} - \gradJ{\tilde{\vtheta}^s}}^2} \nonumber\\
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}
	+\Ets[]{\norm[]{\gradApp{\tilde{\vtheta}^s}{N} - \Ets[]{\gradApp{\tilde{\vtheta}^s}{N}}}^2} \nonumber\\
	&\qquad+ 
	\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i} -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}- \Ets{\score{\vtheta_t^{s+1}}{\tau_i} -
				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}}^2} \nonumber\\
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
	\\
	&\qquad+ 
	\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i} -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}- \Ets{\score{\vtheta_t^{s+1}}{\tau_i} -
				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}}^2} \nonumber\\
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i} -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}^2} \\
	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
	\nonumber\\
	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}
			-\score{\tilde{\vtheta}^s}{\tau_i}
			+\score{\tilde{\vtheta}^s}{\tau_i} 
			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}^2} \nonumber\\
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
	\nonumber\\
	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}
			-\score{\tilde{\vtheta}^s}{\tau_i}}^2} \nonumber
	+\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i} 
			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}^2} \nonumber\\
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
	\nonumber\\
	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i} 
			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}}^2} \\
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
	\nonumber\\
	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+dG^2R^2\Ets{(\omega(\tau_i)-1)^2} \\
	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
	\nonumber\\
	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+dG^2R^2\left(\Ets{\omega^2(\tau_i)}-1\right) \nonumber\\
	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
	\nonumber\\
	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+dG^2R^2\left(M_2-1\right),
\end{align}
where (1) is from the definition of $\gradApp{\vtheta}{N}$, (2) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$, (3) is from L-smoothness, 
(4) is from [todo], and (5) is from [todo].
\end{proof}

>>>>

%%%USEFUL PACKAGES%%%
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{amsmath,amssymb}
\usepackage{mathtools}
\usepackage{pifont}
%%Theorems
\usepackage{amsthm}
\theoremstyle{remark}
\newtheorem*{rem}{Remark}
\newtheorem*{note}{Note}
\newtheorem{case}{Case}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{conj}{Conjecture}
\newtheorem{example}{Example}
%Restatable
\usepackage{thmtools, thm-restate}
\declaretheorem[numberwithin=section]{thm}
\declaretheorem[sibling=thm]{lemma}
\declaretheorem[sibling=thm]{corollary}
\declaretheorem[sibling=thm]{assumption}
\declaretheorem[sibling=thm]{theorem}
%e.g. ...
\usepackage{xspace}
\DeclareRobustCommand{\eg}{e.g.,\@\xspace}
\DeclareRobustCommand{\ie}{i.e.,\@\xspace}
\DeclareRobustCommand{\wrt}{w.r.t.\@\xspace}
%%%%%%

@@

\begin{abstract}

\end{abstract}

\section{Introduction}

\section{Preliminaries}
In this section we provide essential background on policy gradient methods for Reinforcement Learning and stochastic variance-reduced gradient methods for finite sum optimization. In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.

\subsection{Policy Gradient}
A Reinforcement Learning task \cite{sutton2011reinforcement} can be modeled with a discrete-time continuous Markov Decision Process (MDP) $\langle\Sspace,\Aspace,\Transition,\Reward,\gamma,\rho\rangle$, where $\Sspace$ is a continuous state space; $\Aspace$ is a continuous action space; $\Transition$ is a Markovian transition model, where $\Transition(s'\mid s,a)$ defines the transition density from state $s$ to $s'$ under action $a$; $\Reward$ is the reward function, where $\Reward(s,a) \in [-R,R]$ is the expected reward for state-action pair $(s,a)$ and $R$ is the maximum absolute-value reward; $\gamma\in[0,1)$ is the discount factor; and $\rho$ is the initial state distribution.
The agent's behavior is modeled as a policy $\pi$, where $\pi(a\mid s)$ is the density distribution over $\Aspace$ of the action to take in state $s$.
We consider episodic tasks, \ie, tasks composed of episodes of length $H$, also called time horizon. In this setting, a trajectory $\tau$ is the sequence of states and actions $(s_0,a_0,s_1,a_1,\dots,s_{H-1},a_{H-1})$ observed by following some stationary policy in a given episode, where $s_0$ is always sampled from the initial state distribution $\rho$. With abuse of notation, we denote with $\pi(\tau)$ the density distribution induced by policy $\pi$ on the set $\Tspace$ of all possible trajectories, and with $\Reward(\tau)$ the total discounted reward provided by trajectory $\tau$:
\begin{align*}
\Reward(\tau) = \sum_{t=0}^{H-1}\gamma^t\Reward(s_t,a_t).
\end{align*}
Performance $J$ allows to rank policies in terms of expected reward:
\begin{align*}
	J(\pi) = \int_{\Tspace}\pi(\tau)\Reward(\tau)\de \tau.
\end{align*}
Solving an MDP means finding a policy $\pi^*$ maximizing $J$.
\par
Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^m\}$, with the only constraint that $\pol$ is differentiable in the $m$-dimensional parameter vector $\vtheta$. We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we denote $\pol(\tau)$ with $\pi(\tau\mid\vtheta)$ for the sake of readability. The search for a locally optimal policy is performed by means of gradient ascent, where the gradient of $J(\vtheta)$ \wrt the policy parameters is \cite{sutton2000policy}:
\begin{align*}
\gradJ{\vtheta} = \int_{\Tspace}\pol(\tau)\score{}{\tau}\Reward(\tau)\de \tau.
\end{align*}
We consider the on-line learning problem, where trajectories must be sampled directly by interacting with the environment. In this setting, stochastic gradient descent is typically employed: after sampling a trajectory $\tau$ or a batch of $N$ trajectories $\Dataset_{N}$,
policy parameters are updated as $\vtheta\gets\vtheta + \alpha\gradApp{\vtheta}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of the true gradient. The REINFORCE gradient estimator \cite{williams1992simple} provides a simple, unbiased way of estimating the gradient:
\begin{align*}
\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\left(\sum_{h=0}^{H-1}\score{}{a_h^n\mid s_h^n}\right)\left(\sum_{h=0}^{H-1}\gamma^h r_h^n - b\right),
\end{align*}
where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and b can be any baseline, provided it is constant \wrt to actions.
The G(PO)MDP gradient estimator \cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance while preserving the unbiasedness:
\begin{align*}
\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_h^n\mid s_h^n}\right)\left(\gamma^h r_h^n - b\right).
\end{align*}

@@

\section{Convergence Guarantees}
\begin{assumption}\label{ass:bounded_score}
For each trajectory $\tau$, value of $\vtheta$ and gradient component:
\[
	|\score{\vtheta}{\tau}| \leq G
\]
\end{assumption}

\begin{assumption}\label{ass:M2}
	There is a constant $M_2<\infty$ such that, for each pair of policies encountered in Algorithm \ref{alg:svrpg} and for each possible trajectory,
\[
	\mathbb{V}ar\left(\omega(\tau)\right) \leq (M_2-1)
\]
\end{assumption}

\begin{restatable}[]{lemma}{L-smoothness}\label{lemma:lsmooth}
Under Assumption \ref{ass:bounded_score}, $J(\vtheta)$ is L-smooth with Lipschitz constant:
\[
	L = \frac{2RG^2}{1-\gamma}
\]
\end{restatable}

\begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
Under Assumptions \ref{ass:bounded_score} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Algorithm \ref{alg:svrpg} after $T=m\times S$ iterations has, for some constants $\psi>0, D>0$ and $F>0$ and for proper choice of step sizes $\alpha_t$, the following property:
\begin{align*}
	&\EVV[]
	{\norm[]{\nabla J(\vtheta_A)}^2} 
		\leq
		\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T} \\
		&\qquad+
		\frac{D\max_s\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}}{\psi N}
		+\frac{F(M_2-1)}{\psi}
\end{align*}
\end{restatable}

@@

%Lemma 2
\begin{restatable}[]{lemma}{auxtwo}\label{lemma:aux2}
Under Assumption \ref{ass:bounded_score}
, the expected squared norm of the SVRPG gradient can be bounded as follows:
\[
\Ets{\gradBlack{\vtheta_t^{s+1}}} \leq
\Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
\nonumber 
+dG^2R^2\left(M_2-1\right)
\]
\end{restatable}
\begin{proof}

	\begin{align}
	\Ets{\gradBlack{\vtheta_t^{s+1}}} 
	&= \Ets{\norm[]{\gradApp{\tilde{\vtheta}^s}{N}
			+ \score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
	&= \Ets{\norm[]{\gradApp{\tilde{\vtheta}^s}{N}
			+ \score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)-\gradJ{\vtheta_t^{s+1}} + \gradJ{\tilde{\vtheta}^s}
			+\gradJ{\vtheta_t^{s+1}} - \gradJ{\tilde{\vtheta}^s}}^2} \nonumber\\
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2}
	+\Ets[]{\norm[]{\gradApp{\tilde{\vtheta}^s}{N} - \Ets[]{\gradApp{\tilde{\vtheta}^s}{N}}}^2} \nonumber\\
	&\qquad+ 
	\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)- \Ets{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}}^2} \nonumber\\
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
	\\
	&\qquad+ 
	\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)- \Ets{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}}^2} \nonumber\\
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} \nonumber\\
	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \\
	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
	\nonumber\\
	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)
			-\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)
			+\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
	\nonumber\\
	&\qquad+\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)
			-\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber
	+\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
	\nonumber\\
	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \\
	&\leq \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
	\nonumber\\
	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+G^2\frac{R^2}{(1-\gamma)^2}\Ets{(\omega(\tau_i)-1)^2} \\
	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
	\nonumber\\
	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+G^2\frac{R^2}{(1-\gamma)^2}\left(\Ets{\omega^2(\tau_i)}-1\right) \nonumber\\
	&= \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} 
	+\frac{1}{N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}
	\nonumber\\
	&\qquad+L^2\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	+G^2\frac{R^2}{(1-\gamma)^2}\left(M_2-1\right),
\end{align}
where (1) is from the definition of $\gradApp{\vtheta}{N}$, (2) is from $\EVV[]{\norm[]{X-\EVV[]{X}}^2}\leq\EVV[]{\norm[]{X}^2}$ for any stochastic variable $X$, (3) is from L-smoothness, 
(4) is from Assumption \ref{ass:bounded_score}, and (5) is from Assumption \ref{ass:M2}.
\end{proof}

%Lemma 0
\begin{restatable}[]{lemma}{auxzero}\label{lemma:aux0}
Under Assumption \ref{ass:bounded_score},
the expected scalar product between $\gradBlack{\vtheta_t^{s+1}}$ and any function $f(\vtheta_t^{s+1})$ which is deterministic for a fixed $\vtheta_t^{s+1}$ can be bounded as follows:
\begin{align*}
\Ets[t]{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} &\geq
\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \\
&\qquad-
\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2}
\end{align*}
\end{restatable}
\begin{proof}
\begin{align}
	\Ets{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}}
	&=
	\Ets{\dotprod{\gradIdeal{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} +
	\Ets[t-1]{\dotprod{e_s}{f(\vtheta_t^{s+1})}} \\
	&=
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} +
	\Ets[t-1]{\dotprod{e_s}{f(\vtheta_t^{s+1})}} \\
	&=
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad+
	\dotprod{\Ets{e_s}}{\Ets{f(\vtheta_t^{s+1})}}
	+\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}  \nonumber\\
	&= 
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad+
	\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})} \\
	&\geq  
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad-
	\Tr\left|\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{f(\vtheta_t^{s+1})}\right| \nonumber \\
	%
	&\geq
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad-\Tr\left(\sqrt{\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}}}\circ\sqrt{\Varts[t-1]{f(\vtheta_t^{s+1})}}\right) \nonumber\\
	%
	&\geq	
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad-
	\frac{\alpha_t}{2}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{N}} -\frac{\alpha_t}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})}\\
	%
	&=
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad-
	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{1}{2}\Tr\Varts[t-1]{f(\vtheta_t^{s+1})} \\
	%
	&\geq
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{f(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad-
	\frac{1}{2N}\Tr\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}} -\frac{1}{2}\Ets[t-1]{\norm[]{f(\vtheta_t^{s+1})}^2},\nonumber
\end{align}
where (6) is from Definition \ref{def:ideal}; (7) is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $f(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$; (8) is from $\Ets{e_s}=0$; (9) is from Young's inequality; and (10) is from the definition of $\gradApp{\vtheta}{N}$.
\end{proof}
